Planetary perturbers, or: flaring star-planet interactions in Kepler and TESS

### 2023-06-23 src/tex/ms.tex
[showyourwork] overleaf sync

% Title
\title{Flaring star-planet interactions in Kepler and TESS}

@@

    In the 1811 searched systems, we found 25 single stars with three or more flares each. We quantified the significance of the clustering in each system, and compared it against the theoretically expected amplitude of magnetic interaction that leads to planet-induced flaring. Most systems do not show any clustering, as expected from their relatively wide orbits. Those we expect to show planet-induced flaring fall on two branches. An inactive one, without any signs of clustering, and a tentative active one, where the clustering becomes more pronounced as the expected amplitude of interaction increases. The flares in HIP 67522, a young Hot Jupiter system, are prominently clustered ($p<0.006$), identifying it as the most promising candidate for magnetic star-planet interaction in our sample.

@@

On the one hand, a close-in planet that changes the star's rotation and activity makes it more difficult for us to model the system's evolution, as it affects fundamental stellar parameters. Ages inferred from rotation will be biased, if the star is spun up or spun down relative to stars without close-in companions~\citep{tejadaarevalo2021further, brown2014discrepancies, maxted2015comparison}. Analogously, altered activity levels may mimic an older or younger host than it actually is~\citep{ilic2022tidal}. On the other hand, measuring in what ways and how much close-in planets alter their hosts' behavior can provide important insights into the habitability of the entire system~\citep{shkolnik2018signatures}. Even if they orbit far inside the habitable zone, such planets probe the system's space weather and mass loss~~\citep{cohen2011dynamics, cohen2015interaction, hazra2022impact} by revealing the extent of the Alfv\'en zone through SPI~\citep{kavanagh2021planetinduced, chebly2022destination}. % Prospectively, planet-induced stellar activity could quantify otherwise difficult to measure planetary properties such as the strength and extent of their magnetic field~\citep{cauley2019magnetic}.

The search for magnetic SPI (Section~\ref{sec:intro:mspi}) in the form of changes in global activity indicators~(Section~\ref{sec:intro:global}), has so far been a series of detections and non-detections, often due to selection bias in the studied systems. Looking for local changes that trace the magnetic footpoint that connects planet to star mitigates this bias, but introduces another -- sampling bias due to poor orbital phase coverage. Flares~(Section~\ref{sec:intro:flares}) are activity markers that can be caused by SPI~(Section~\ref{sec:intro:fspi}). They can be monitored with high phase coverage over long periods of time with missions like Kepler and TESS. Thanks to their large archives of hundreds of cumulative years of stellar monitoring, we can overcome the limitations of individual system studies~(Section~\ref{sec:intro:local}), and now look for flaring SPI more comprehensively, and systematically.

@@

\subsection{Searching for global changes in activity}
\label{sec:intro:global}
Measuring magnetic star planet-interaction is notoriously difficult in individual systems because a non-detection may not mean an absence of the interaction, but merely a temporary cessation~\citep{shkolnik2005hot, shkolnik2008nature}. Statistical studies searching for changes in global chromospheric and coronal activity indicators in star-planet systems with close-in planets relative to planet-less stars are inconclusive~\citep{kashyap2008extrasolar,scharf2010possible, shkolnik2013ultraviolet, france2018farultraviolet, viswanath2020statistical, krejcova2012evidence, miller2015comprehensive, poppenhaeger2010coronal}. In these studies, it is challenging both to define a control sample to measure the activity against, and to quantify how the way we detect planets introduces selection bias in the activity measure. It is, for instance, easier to detect a planet around a magnetically inactive star, because both radial velocity and transit detections will be less affected by stellar variability~\citep{poppenhaeger2011correlation}. Tidal interactions can also induce global increase in activity~\citep{ilic2022tidal}, which adds further ambiguity to the interpretation.

\subsection{Flares}
\label{sec:intro:flares}
In this study, we will look at stellar flares as another way of dissipating the energy stored in the magnetic field. Flares are strong and impulsive eruptions released by reconnection and subsequent relaxation of field lines in the corona~\citep{svestka1976solar,priest2002magnetic}. In contrast to solar flares, stellar flares often are extremely energetic, sometimes enhancing the stellar flux by orders of magnitude~\citep{maehara2012superflares, shibayama2013superflares, paudel2018k2} so that they can be detected in stars over 1kpc away~\citep{chang2015photometric}. But since they occur randomly in time, individual events can be hard to catch. %We can observe flares in nearly all electromagnetic bands -- from radio to X-ray~\citep{benz2010physical}.

@@

If we picture the magnetic field lines connecting the planet to the star in the sub-Alfv\'enic zone, we expect that the stellar footpoint of these lines will move along with the planet. Flares triggered by the planet will occur at a preferred orbital phase, namely whenever the footpoint is visible to the observer. Depending on the system's viewing geometry, i.e., the inclination of the rotation axis, the spin-orbit (mis-)alignment, and the extent and configuration of the large scale magnetic field -- the footpoint could be always in view, or never. In expectation, however, in star-planet systems with favorable viewing geometries and magnetic field strengths and configurations, the footpoint visibility will be modulated in phase with the planetary orbit~(Fig.~\ref{fig:sketch}). The resulting phase preference of planet-induced flares introduces a deviation from a random distribution of flares. Space missions like Kepler and TESS provide light curves that cover all phases of the planet's orbit, so that phase-correlated planet-induced flares can be detected against a background of uncorrelated intrinsic flaring.

@@

\subsection{Overview}
In this work, we searched for excess flaring in all available Kepler and TESS short cadence light curves of confirmed star-planet systems~(Section~\ref{sec:data}). We scoured the light curves for flares, and tested their observed phases for departures from a uniform distribution in phase with the innermost planet's orbit~(Section~\ref{sec:methods}). We present the catalog of all flares found in our sample of star-planet systems, and compare the test results with the expected power of interaction for each system~(Section~\ref{sec:results}). We discuss the results for the most interesting star-planet systems individually~(Section~\ref{sec:results:individualstars}), and take a look at the patterns emerging from our analysis~(Section~\ref{sec:discussion}), and the possible role of tidal flaring SPI in our observations~(Section~\ref{sec:discussion:tidal}).  We summarize our findings in Section~\ref{sec:summary}.

@@

\section{Data}
\label{sec:data}
In this study, we aimed at compiling the largest possible sample to search for flaring SPI. We took the entire Planetary Systems Composite Parameters Table (PSCP\footnote{ \url{https://exoplanetarchive.ipac.caltech.edu/cgi-bin/TblView/nph-tblView?app=ExoTbls&config=PSCompPars}}, Section~\ref{sec:data:sps}) from the NASA Exoplanet Archive, as of July 2022, and scanned the systems for available Kepler and TESS light curves~(Section~\ref{sec:data:photometry}). We searched a total of 1811 systems and over 7200 light curves from both missions for flares. To calculate the amount of excess flaring from SPI, we further required the orbital periods of each system~(Section~\ref{sec:data:orbitalperiod}). To compare our measurements to theory, we estimated the expected power of interaction for each system with at least three flares. For this, we required further system parameters -- distance between planet and star, stellar rotation period, planetary radius, and stellar luminosity~(Sections~\ref{sec:data:a}-\ref{sec:data:lum}), which we mostly took from the literature. We began by compiling the table of system parameters with the PSCP, but verified, supplemented and updated the table with more up-to-date results. The parameters for the systems in the final sample of 25 systems are shown in Table~\ref{tab:maintable_lit}.

\begin{table*}
\movetableright=-20mm
\footnotesize
    \script{paper_main_table.py}
    \caption{Flaring single star-planet system parameters.}
    \input{output/table_lit_vals.tex}
        \label{tab:maintable_lit}
    \tablerefs{\input{output/lit_table_bibstring.tex}}
\end{table*}

We note that we found K2-354 under EPIC 211314705 and K2-329(!) in \cite{bouma2020cluster}, who refer to the detection paper \cite{pope2016transiting}, and also under TIC 468989066.

@@

The Transiting Exoplanet Survey Satellite~[TESS, ][]\citep{ricker2015transiting} is an all-sky mission that began operations in 2018, completed two sky scans until summer 2022. It is observing at the time of writing, collecting nearly continuous photometric time series in the $600-1000\,$nm band for $\sim 27\,$d in each observing Sector. About $200\,000$ stars have been observed in 2-min cadence in the first two years of operations alone, with about $20,000$ targets per Sector. Out of these, from Sector 27 on, $1,000$ targets were observed at even higher 20-s cadence in each Sector. 
% 20000 per sector does not add up, because of stars observed in multiple sectors, the 20000 and 1000 figures can be found here: https://tess.mit.edu/observations/target-lists/
% extended TESS mission https://heasarc.gsfc.nasa.gov/docs/tess/extended.html

Based on the filtered PSCP table, we use the \texttt{lightkurve}~\citep{lightkurvecollaboration2018lightkurve} Python software to query the full Kepler archive (quarters Q0-Q17, Data Release 25), and the most recent TESS catalog (July 2022) for their respective 1-min and 2-min/20-s cadence light curves. In total, we searched 7258 light curves for flares -- 3030 Kepler Quarters, and 4228 TESS Sectors. We found and analyzed light curves for a total of 1811 systems, 344 of which were observed only by the primary Kepler mission, 1205 only by TESS, and 262 by both missions. We did not use K2 light curves because of the many systematics, and consequently the high time investment of de-trending them. But we included the flare list obtained by~\citet{paudel2018k2} for TRAPPIST-1 K2 flares, since this system is a potential candidate for sub-Alfv\'enic interactions~\citep{fischer2019timevariable}.

@@

If the eccentricity is known, then the uncertainty on the distance is set to either the error on the semi-major axis, or to half of the difference between periastron and apoastron -- whichever is larger. If eccentricity is not known, then the uncertainty on the distance is set to either the 25\% error on the semi-major axis, that is, assuming $e=0.25$, or the uncertainty on the semi-major axis -- whichever is larger. We chose $e=0.25$ because it is both a typical value within our sample~(see Table~\ref{tab:maintable_lit}), and in the literature~\citep{eylen2019orbital}.

@@

\subsection{Planetary radii}
\label{sec:data:planetradii}
For the transiting planets, we adopt the literature values for planet radius $R_p$, and uncertainties, from the PSCP. For the non-transiting radial velocity detected planets we use the planetary mass $M_p$ or $M_p\sin i$ to calculate the radius or a lower limit for the radius using the empirical relations derived in \cite{chen2017probabilistic} using their open source \texttt{forecaster} tool, upgraded to \texttt{astro-forecaster}\footnote{https://pypi.org/project/astro-forecaster/} by Ben Cassesse. We note that for GJ 674, the mass estimates in \cite{bonfils2007harps} and \cite{boisse2011disentangling} do not quote uncertainties, but differ by $0.3M_\oplus$, so we assumed that value as the uncertainty on $M_p\sin i$.

@@

\section{Methods}
\label{sec:methods}
We measure flaring SPI as the presence of excess flares triggered by the orbiting planet. In the absence of flaring SPI, flare peak times will be distributed randomly in orbital phase. In the presence of flaring SPI, we measure a phase dependent deviation from this randomness.

@@

We vet all flare candidates by eye, and exclude instrumental false positives and physical ones like passing Solar System Objects. For each confirmed flare candidate, we calculate the relative amplitude $a$ and equivalent duration $ED$, defined as the integrated flare flux $F_{\rm flare}$ over the duration of the flare, divided by the median flux $F_0$ of the star, integrated over the flare duration~\citep{gershberg1972results}:
\begin{equation}
\label{eq:ED}
ED=\displaystyle \int \mathrm dt\, \frac{F_{\rm flare}(t)}{F_0}.
\end{equation}
Equivalent duration is equal to the time during which the non-flaring star releases as much energy as the flare.

\subsection{Custom Anderson-Darling test}
\label{sec:methods:adtest}
For each star-planet system with three or more flares in its light curves, we test for deviations from a random distribution of flares with orbital phase. We use the same customized Anderson-Darling test as in \cite{ilin2022searching}, which is more sensitive to deviations at both ends of the distribution than the widely used Kolmogorov-Smirnov test~\citep{feigelson2012modern}. In brief, we first take the number of flares observed in the TESS and Kepler data to calculate a base flare rate for each light curve. We then calculate how often each orbital phase has been observed, also for each light curve separately, to ensure that the different detection thresholds for TESS and Kepler do not bias the base rate. With the phase coverage and base flare rate combined, we can tell how many flares we would expect to see in any given phase bin if flares were randomly distributed. We then aggregate this number of expected flares per time bin into one expected distribution by summing over all available light curves. As a last step, we compare the expected distribution to the observed one using an Anderson-Darling test. We repeat this test with different phase offsets to account for potential biases in sensitivity of the test at different phases.

@@

Our goal was to measure magnetic SPI as the statistical clustering of flares in phase with the innermost planet's orbit. We searched all star-planet systems that were observed with Kepler and TESS at 1 or 2 min cadence, respectively, for flares. For the star-planet systems in the resulting flare catalog~(Section~\ref{sec:results:catalog}), we confirmed that the orbital phase could be known well enough for the entire observing baseline of the system~(Section~\ref{sec:results:coherence}). We calculated the orbital phases of each flare, and the flare phase distributions for each system~(Section \ref{sec:results:phasedist}). We then applied the custom Anderson-Darling test to assess how much each distribution was different from randomly distributed intrinsic flaring. Comparing the test results with the expected power of magnetic SPI, we found our main result: The clustering of flares in orbital phase tentatively increases with the expected power in all considered scenarios of magnetic SPI~(Section~\ref{sec:results:spi}). However, not all systems with high expected power show signs of flaring SPI, which creates two branches in the distribution of test results. Finally, in Section~\ref{sec:results:individualstars}, we provide additional context for the most interesting systems in our sample, and explain why we had to exclude others from the analysis.

\subsection{Flare catalog}
\label{sec:results:catalog}
We searched a total of 3032 Kepler Quarters and 4181 TESS Sectors of a total of 1811 star-planet systems for flares. We inspected all candidates manually, and added flares observed by K2 on TRAPPIST-1~\citep{paudel2018k2}, and flares from the planet hosting primary in the Kepler-411 binary~\citep{jackman2021stellara}. The final table contains a total of \input{output/PAPER_total_number_of_flares.txt}flares in \input{output/PAPER_total_number_of_systems_with_flares.txt}systems. In Table~\ref{tab:flares}, we list the flares and their characteristics. To identify each event, we provide the name of the system in which the flare was found, along with its TIC, whether it was observed with Kepler or TESS, and during what Quarter or Sector, respectively. For each flare, we also give the start and finish time, relative amplitude $a$ and equivalent duration $ED$. If the transit midtime is known for the innermost planet, we also give the orbital phase at which the flare occurred. The full table is provided with the supplementary material, and via Zenodo (see Data Availability Statement).

@@

\begin{equation}
    \tau = P^2 / \sigma_P,
\end{equation}
where $P$ is the orbital period, and $\sigma_P$ is its uncertainty.
If the orbital period is well-known, the resulting ratio $\Delta T/\tau$ should be $<<1$. The bottom panel in Figure~\ref{fig:coherence_hist} shows a histogram of $\Delta T/\tau$ in our sample. The orbital period is known sufficiently well for our analysis in most cases with ratios $<0.02$, that is, the orbital phase of the last flare is at most $2\%$ off, compared to the first. The three cases where the ratio is highest with $0.05$, $0.12$, $0.24$ deviation are TAP 26, GJ 3082, and GJ 674, respectively. We keep these systems in our sample, but caution that the measured absence of flaring SPI might be due to the uncertain orbital period of the planet in these cases.

@@

Figure~\ref{fig:adtest_bp} shows that in every scenario we considered (Alfv\'en wing or stretch-and-break mechanism, with magnetized or unmagnetized planets), there are two branches on the high end of expected powers. On one branch, the significance of the measured interaction increases with the expected power $P_{\rm spi}$. On the other branch, no interaction is measured regardless of $P_{\rm spi}$.

Table~\ref{tab:maintable_der} lists the $p$-values for each star-planet system. There is no star-planet system for which we find a $>3\sigma$ ($p<0.0027$) signal of flaring SPI in our sample. However, overall, the significance of the deviation increases with higher expected power $P_{\rm spi}$ of interaction, where additional subscripts denote different scenarios introduced in Section~\ref{sec:methods:pspi}. Table~\ref{tab:maintable_der} also lists the derived parameters required in these scaling laws, i.e., Rossby number $R$o, surface-average magnetic field strength $B$, and relative velocity between the planet and a co-rotating magnetic field $v_{\rm rel}$.

@@

\begin{table*}
\footnotesize
\movetableright=-20mm
\script{paper_main_table.py}
\caption{Flaring star-planet interaction. $Ro$, $B$, and $v_{rel}$, are derived from literature values (Table~\ref{tab:maintable_lit}). $P_{\rm spi,xx}$ stands for the power of stretch-and-break ($sb$) and Alfv\'en wing ($aw$) interaction mechanisms, assuming the planet has a magnetic field strength of $1\,$G. $P_{\rm spi,xx0}$ is the same, but assuming an unmagnetized planet. All powers are normalized to AU Mic. The $p$-value of the Anderson-Darling test is lower when the system shows more flares periodic with the planetary orbit.}
\input{output/table_der_vals.tex}
    \label{tab:maintable_der}

@@

\begin{figure*}[ht!]
    \script{paper_adtest_vs_value_scatterplots.py}
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/PAPER_ADtest_bg_main.png}
        \caption{
            Expected power of magnetic SPI vs. AD test results, assuming the stretch-and-break scenario. While the distribution of $p$-values is consistent with no interaction, all scenarios indicate lower $p$-values only for high expected powers of interaction. See Fig.~\ref{fig:adtest_ro} in the Appendix for the same figure, color-coded by Rossby number.
        }
        \label{fig:adtest_bp}
    \end{centering}
\end{figure*}

>>>>

% Title
\title{Planetary perturbers, or: flaring star-planet interactions in Kepler and TESS}

@@

    In the 1811 searched systems, we found 25 single stars with three or more flares each. We quantified the significance of the clustering in each system, and compared it against the theoretically expected power of magnetic interaction that leads to planet-induced flaring. Most systems do not show any clustering, as expected from their relatively wide orbits. Those we expect to show planet-induced flaring fall on two branches. An inactive one, without any signs of clustering, and a tentative active one, where the clustering becomes more pronounced as the expected power of interaction increases. The flares in HIP 67522, a young Hot Jupiter system, are prominently clustered ($p<0.006$), identifying it as the most promising candidate for magnetic star-planet interaction in our sample.

@@

On the one hand, a close-in planet that changes the star's rotation and activity makes it more difficult for us to model the system's evolution, as it affects fundamental stellar parameters. Ages inferred from rotation will be biased, if the star is spun up or down relative to stars without close-in companions~\citep{tejadaarevalo2021further, brown2014discrepancies, maxted2015comparison}. Analogously, altered activity levels may mimic an older or younger host than it actually is~\citep{ilic2022tidal}. On the other hand, measuring in what ways and how much close-in planets alter their hosts' behavior can provide important insights into the habitability of the entire system~\citep{shkolnik2018signatures}. Even if they orbit far inside the habitable zone, such planets probe the system's space weather and mass loss~~\citep{cohen2011dynamics, cohen2015interaction, hazra2022impact} by revealing the extent of the Alfv\'en zone through SPI~\citep{kavanagh2021planetinduced, chebly2022destination}. % Prospectively, planet-induced stellar activity could quantify otherwise difficult to measure planetary properties such as the strength and extent of their magnetic field~\citep{cauley2019magnetic}.

The search for magnetic SPI (Section~\ref{sec:intro:mspi}) in the form of changes in global activity indicators~(Section~\ref{sec:intro:global}) has so far been a series of detections and non-detections, often due to selection bias in the studied systems. Looking for local changes that trace the magnetic footpoint that connects planet to star mitigates this bias, but introduces another -- sampling bias due to poor orbital phase coverage. Flares~(Section~\ref{sec:intro:flares}) are activity markers that can be caused by SPI~(Section~\ref{sec:intro:fspi}). They can be monitored with high phase coverage over long periods of time with missions like Kepler and TESS. Thanks to their large archives of hundreds of cumulative years of stellar monitoring, we can overcome the limitations of individual system studies~(Section~\ref{sec:intro:local}), and now look for flaring SPI more comprehensively, and systematically.

@@

\subsection{Searching for global changes in activity}
\label{sec:intro:global}
Measuring magnetic star planet-interaction is notoriously difficult in individual systems because a non-detection may not mean an absence of the interaction, but merely a temporary cessation~\citep{shkolnik2005hot, shkolnik2008nature}. Statistical studies searching for changes in global chromospheric and coronal activity indicators in star-planet systems with close-in planets relative to planet-less stars are inconclusive~\citep{kashyap2008extrasolar,scharf2010possible, shkolnik2013ultraviolet, france2018farultraviolet, viswanath2020statistical, krejcova2012evidence, miller2015comprehensive, poppenhaeger2010coronal}. In these studies, it is challenging both to define a control sample to measure the activity against, and to quantify how the way we detect planets introduces selection bias in the activity measure. It is, for instance, easier to detect a planet around a magnetically inactive star, because both radial velocity and transit detections will be less affected by stellar variability~\citep{poppenhaeger2011correlation}, but that does not imply that SPI quenches activity. Further, tidal interactions can induce global increase in activity~\citep{ilic2022tidal}, which adds a physical ambiguity to the interpretation~(see also Section~\ref{sec:discussion:tidal}).

\subsection{Flares}
\label{sec:intro:flares}
In this study, we look at stellar flares as another way of dissipating the energy stored in the magnetic field. Flares are strong and impulsive eruptions released by reconnection and subsequent relaxation of field lines in the corona~\citep{svestka1976solar,priest2002magnetic}. In contrast to solar flares, stellar flares often are extremely energetic, sometimes enhancing the stellar flux by orders of magnitude~\citep{maehara2012superflares, shibayama2013superflares, paudel2018k2} so that they can be detected in stars over 1kpc away~\citep{chang2015photometric}. But since they occur randomly in time, individual events can be hard to catch. %We can observe flares in nearly all electromagnetic bands -- from radio to X-ray~\citep{benz2010physical}.

@@

If we picture the magnetic field lines connecting the planet to the star in the sub-Alfv\'enic zone, we expect that the stellar footpoint of these lines will move along with the planet. Flares triggered by the planet will occur at a preferred orbital phase, namely whenever the footpoint is visible to the observer. Depending on the system's viewing geometry, i.e., the inclination of the rotation axis, the spin-orbit (mis-)alignment, and the extent and configuration of the large scale magnetic field -- the footpoint could be always in view, or never. In expectation, however, in star-planet systems with favorable viewing geometries and magnetic field strengths and configurations, the footpoint visibility will be modulated in phase with the planetary orbit~(Fig.~\ref{fig:sketch}). The resulting phase preference of planet-induced flares introduces a deviation from the otherwise random distribution of flares~\citep[see, e.g.,][]{doyle2018investigating,feinstein2020flare,howard2021evryflare}. Space missions like Kepler and TESS provide light curves that cover all phases of the planet's orbit, so that phase-correlated planet-induced flares can be detected against a background of uncorrelated intrinsic flaring.

@@

\subsection{Overview}
In this work, we searched for excess flaring in all available Kepler and TESS short cadence light curves of confirmed star-planet systems~(Section~\ref{sec:data}). We scoured the light curves for flares, and tested their observed phases for departures from a uniform distribution in phase with the innermost planet's orbit~(Section~\ref{sec:methods}). We present the catalog of all flares found in our sample of star-planet systems, and compare the test results with the expected power of interaction for each system~(Section~\ref{sec:results}). We discuss the results for the most interesting star-planet systems individually~(Section~\ref{sec:results:individualstars}), and take a look at the patterns emerging from our analysis~(Section~\ref{sec:discussion}), as well as the possible role of tidal flaring SPI in our observations~(Section~\ref{sec:discussion:tidal}).  We summarize our findings in Section~\ref{sec:summary}.

@@

\section{Data}
\label{sec:data}
In this study, we aimed at compiling the largest possible sample to search for flaring SPI. We took the entire Planetary Systems Composite Parameters Table (PSCP\footnote{ \url{https://exoplanetarchive.ipac.caltech.edu/cgi-bin/TblView/nph-tblView?app=ExoTbls&config=PSCompPars}}, Section~\ref{sec:data:sps}) from the NASA Exoplanet Archive, as of July 2022, and scanned the systems for available Kepler and TESS light curves~(Section~\ref{sec:data:photometry}). We searched a total of 1811 systems and over 7200 light curves from both missions for flares. To calculate the clustering of in orbital phase, we further required the orbital periods of each system~(Section~\ref{sec:data:orbitalperiod}). We then estimated the theoretically expected power of interaction for each system with at least three flares. For this, we required further system parameters -- distance between planet and star, stellar rotation period, planetary radius, and stellar luminosity~(Sections~\ref{sec:data:a}-\ref{sec:data:lum}), which we mostly took from the literature. We began by compiling the table of system parameters with the PSCP, but verified, supplemented and updated the table with more up-to-date measurements. The relevant properties of the final sample of 25 systems are shown in Table~\ref{tab:maintable_lit}.

\begin{table*}
\movetableright=-20mm

    \script{paper_main_table.py}
    \caption{Flaring single star-planet system parameters.}

    \input{output/table_lit_vals.tex}
        \label{tab:maintable_lit}
    \tablerefs{\input{output/lit_table_bibstring.tex}}
\end{table*}

%We note that we found K2-354 under EPIC 211314705 and K2-329(!) in \cite{bouma2020cluster}, who refer to the detection paper \cite{pope2016transiting}, and also under TIC 468989066.

@@

The Transiting Exoplanet Survey Satellite~\citep[TESS,][]{ricker2015transiting} is an all-sky mission that began operations in 2018, completed two sky scans until summer 2022. It is observing at the time of writing, collecting nearly continuous photometric time series in the $600-1000\,$nm band for $\sim 27\,$d in each observing Sector. About $200\,000$ stars have been observed in 2-min cadence in the first two years of operations alone, with about $20,000$ targets per Sector. Out of these, from Sector 27 on, $1,000$ targets were observed at even higher 20-s cadence in each Sector. 
% 20000 per sector does not add up, because of stars observed in multiple sectors, the 20000 and 1000 figures can be found here: https://tess.mit.edu/observations/target-lists/
% extended TESS mission https://heasarc.gsfc.nasa.gov/docs/tess/extended.html

Based on the filtered PSCP table, we used the \texttt{lightkurve}~\citep{lightkurvecollaboration2018lightkurve} Python software to query the full Kepler archive (quarters Q0-Q17, Data Release 25), and the most recent TESS catalog (July 2022) for their respective 1-min and 2-min/20-s cadence light curves. In total, we searched 7213 light curves for flares -- 3032 Kepler Quarters, and 4181 TESS Sectors. We found and analyzed light curves for a total of 1811 systems, 344 of which were observed only by the primary Kepler mission, 1205 only by TESS, and 262 by both missions. We did not use K2 light curves because of the many systematics, and consequently the high time investment of de-trending them. But we included the flare list obtained by~\citet{paudel2018k2} for TRAPPIST-1 K2 flares, since this system is a potential candidate for sub-Alfv\'enic interactions~\citep{fischer2019timevariable}.

@@

If the eccentricity is known, then the uncertainty on the distance is set to either the error on the semi-major axis, or to half of the difference between periastron and apoastron -- whichever is larger. If eccentricity is not known, then the uncertainty on the distance is set to either a 25\% error on the semi-major axis, that is, assuming $e=0.25$, or the uncertainty on the semi-major axis -- whichever is larger. We chose $e=0.25$ because it is both a typical value within our sample~(see Table~\ref{tab:maintable_lit}), and in the literature~\citep{eylen2019orbital}.

@@

\subsection{Planetary radii}
\label{sec:data:planetradii}
For the transiting planets, we adopt the literature values for planet radius $R_p$, and uncertainties, from the PSCP. For the non-transiting radial velocity detected planets we use  $M_p\sin i$ to calculate a lower limit for the radius using the empirical relations derived in \cite{chen2017probabilistic} using their open source \texttt{forecaster} tool, upgraded to \texttt{astro-forecaster}\footnote{https://pypi.org/project/astro-forecaster/} by Ben Cassesse. We note that for GJ 674, the mass estimates in \cite{bonfils2007harps} and \cite{boisse2011disentangling} do not quote uncertainties, but differ by $0.3M_\oplus$, so we assumed that value as the uncertainty on $M_p\sin i$.

@@

\section{Methods}
\label{sec:methods}
We measure flaring SPI as the presence of flares triggered by the orbiting planet. In the absence of flaring SPI, flare peak times will be distributed randomly in orbital phase. In the presence of flaring SPI, we measure a phase dependent deviation from this randomness.

@@

We vet all flare candidates by eye, and exclude instrumental and physical false positives, such as Solar System Objects passing by the field of view. For each confirmed flare candidate, we calculate the relative amplitude $a$ and equivalent duration $ED$, defined as the integrated flare flux $F_{\rm flare}$ over the duration of the flare, divided by the median flux $F_0$ of the star, integrated over the flare duration~\citep{gershberg1972results}:
\begin{equation}
\label{eq:ED}
ED=\displaystyle \int \mathrm dt\, \frac{F_{\rm flare}(t)}{F_0}.
\end{equation}
Equivalent duration is equal to the time during which the non-flaring star releases as much energy as the flare.

\subsection{Custom Anderson-Darling test}
\label{sec:methods:adtest}
For each star-planet system with three or more flares in its light curves, we test for deviations from a random distribution of flares with orbital phase. We use the same customized Anderson-Darling test as in \cite{ilin2022searching}, which is more sensitive to deviations at both ends of the distribution than the widely used Kolmogorov-Smirnov test~\citep{feigelson2012modern}. In brief, we first take the number of flares observed in the TESS and Kepler data to calculate a base flare rate for each light curve. We then calculate how often each orbital phase has been observed, also for each light curve separately, to ensure that the different detection thresholds for TESS and Kepler do not bias the base rate. With the phase coverage and base flare rate combined, we can tell how many flares we would expect to see in any given phase bin if flares were randomly distributed. We then aggregate this number of expected flares per time bin into one expected distribution by summing over all available light curves. As a last step, we compare the expected distribution to the observed one using an Anderson-Darling test. We repeat this test with different phase offsets to account for remaining biases in sensitivity of the test at different phases.

@@

Our goal was to measure magnetic SPI as the statistical clustering of flares in phase with the innermost planet's orbit. We searched all star-planet systems that were observed with Kepler and TESS at 1 or 2 min cadence, respectively, for flares. For the star-planet systems in the resulting flare catalog~(Section~\ref{sec:results:catalog}), we confirmed that the orbital phase could be known well enough for the entire observing baseline of the system~(Section~\ref{sec:results:coherence}). We calculated the orbital phases of each flare, and the flare phase distributions for each system~(Section \ref{sec:results:phasedist}). We then applied a custom Anderson-Darling test~(see Section~\ref{sec:methods:adtest}) to assess how much each distribution was different from randomly distributed intrinsic flaring. Comparing the test results with the expected power of magnetic SPI, we found our main result: The clustering of flares in orbital phase tentatively increases with the expected power in all considered scenarios of magnetic SPI~(Section~\ref{sec:results:spi}). However, not all systems with high expected power show signs of flaring SPI, which creates two branches in the distribution of test results. Finally, in Section~\ref{sec:results:individualstars}, we provide additional context for the most interesting systems in our sample, and explain why we had to exclude others from the analysis.

\subsection{Flare catalog}
\label{sec:results:catalog}
We searched a total of 3032 Kepler Quarters and 4181 TESS Sectors of a total of 1811 star-planet systems for flares. We inspected all candidates manually, and added flares observed by K2 on TRAPPIST-1~\citep{paudel2018k2}, and flares from the planet hosting primary in the Kepler-411 binary~\citep{jackman2021stellara}. The final table contains a total of \input{output/PAPER_total_number_of_flares.txt}flares in \input{output/PAPER_total_number_of_systems_with_flares.txt}systems. In Table~\ref{tab:flares}, we list the flares and their characteristics. To identify each event, we provide the name of the system in which the flare was found, along with its TIC, whether it was observed with Kepler or TESS, and during what Quarter or Sector, respectively. For each flare, we also give the start and finish time, relative amplitude $a$ and equivalent duration $ED$~(Eq.~\ref{eq:ED}). If the transit midtime is known for the innermost planet, we also give the orbital phase at which the flare occurred. The full table is available online (see Data Availability Statement).

@@

\begin{equation}
    \tau = P^2 / \sigma_P,
\end{equation}
where $P$ is the orbital period, and $\sigma_P$ is its uncertainty.
If the orbital period is well-known, the resulting ratio $\Delta T/\tau$ should be $\ll 1$. The bottom panel in Figure~\ref{fig:coherence_hist} shows a histogram of $\Delta T/\tau$ in our sample. The orbital period is known sufficiently well for our analysis in most cases with ratios $<0.02$, that is, the orbital phase of the last flare is at most $2\%$ off, compared to the first. The three cases where the ratio is highest with $0.05$, $0.12$, $0.24$ deviation are TAP 26, GJ 3082, and GJ 674, respectively. We keep these systems in our sample, but caution that the measured absence of flaring SPI might be partly due to the uncertain orbital period of the planet in these cases.

@@

Figures~\ref{fig:adtest_bp} and \ref{fig:adtest_bp_rest} show that in every scenario we considered (Alfv\'en wing or stretch-and-break mechanism, with magnetized or unmagnetized planets, see Section~\ref{sec:methods:pspi}), there are two branches on the high end of expected powers. On one branch, the significance of the measured interaction increases with the expected power $P_{\rm spi}$. On the other branch, no interaction is measured regardless of $P_{\rm spi}$.

Table~\ref{tab:maintable_der} lists the $p$-values for each star-planet system. There is no star-planet system for which we find a $>3\sigma$ ($p<0.0027$) signal of flaring SPI in our sample. However, overall, the significance of the deviation increases with higher expected power $P_{\rm spi}$ of interaction, introduced in Section~\ref{sec:methods:pspi}. Table~\ref{tab:maintable_der} also lists the derived parameters required in these scaling laws, i.e., Rossby number $R$o, surface-average magnetic field strength $B$, and relative velocity between the planet and a co-rotating magnetic field $v_{\rm rel}$.

@@

\begin{table*}
\footnotesize
\movetableright=-20mm
\script{paper_main_table.py}
\caption{Flaring star-planet interaction. $Ro$, $B$, and $v_{rel}$, are derived from literature values (Table~\ref{tab:maintable_lit}). $P_{\rm spi,xx}$ stands for the power of stretch-and-break ($sb$) and Alfv\'en wing ($aw$) interaction mechanisms, assuming the planet has a magnetic field strength of $1\,$G. $P_{\rm spi,xx0}$ is the same, but assuming an unmagnetized planet. All powers are normalized to AU Mic. The $p$-value of the Anderson-Darling test is lower when the system shows a stronger clustering of flares in orbital phase.}
\input{output/table_der_vals.tex}
    \label{tab:maintable_der}

@@

\begin{figure*}[ht!]
    \script{paper_adtest_vs_value_scatterplots.py}
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/PAPER_ADtest_bg_main.png}
        \caption{
            Expected power of magnetic SPI vs. AD test results, assuming the stretch-and-break scenario with an unmagnetized planet. While no individual system amounts to a $>3\sigma$ detection, the global trend is intriguing: Systems with low expected power show no deviation from random intrinsic flaring (i.e. high $p$-values). In contrast, systems with high expected power show two branches. One where the significance of interaction increases with power, and one without such a trend. See Fig.~\ref{fig:adtest_ro} in the Appendix for the same figure, color-coded by Rossby number.
        }
        \label{fig:adtest_bp}
    \end{centering}
\end{figure*}
